---
title: "Assignment3"
author: "Rabin Paudel"
date: "9/24/2021"
output: pdf_document
---

## Problem 1
#### We can instill this library and load the dataset using the 
#### folowing code chunk.

```{r}
library("dslabs")
#install.packages("dplyr")
library("dplyr")
library("tidyr")
admissions
```

#### Next use unite (also from the tidyr library) to combine the columns for 
#### name and gender into one column

```{r}

admissions_new <- admissions %>% 
  pivot_longer(cols = c("admitted","applicants"), 
                                        names_to = "name", values_to = "value") 
admissions_new
```

#### Finally use pivot_wider from the tidyr library to get the required 
#### data frame.

```{r}
New_value <- admissions_new %>% 
  unite(rate,name,gender)
New_value %>% 
  pivot_wider(names_from = "rate")
```

## Problem 2
#### This problem is from Chapter 5 of the book Modern Data Science with R. 
#### The problem uses the Batting, Pitching, and Master data frames in the 
#### Lahman package.


```{r}
#install.packages("Lahman")
library(Lahman)
#Batting
#Pitching
#Master
```
### First Part
### Using the above data frames, answer the following questions.
#### Name every player in baseball history who has accumulated at least 300 home
#### runs (HR column) and at least 300 stolen bases (SB column). You can find 
#### the first and last name of the player in the Master data frame. Join this 
#### to your result along with the total home runs and total bases stolen for 
#### each of these elite players.

```{r}
mas <- Master %>% 
  group_by(playerID) %>%
  summarize( playerID = paste(playerID),
             name = paste(nameFirst, nameLast, sep = " "))

fg <- Batting %>% 
  group_by( playerID) %>% 
  summarize(HR = sum(HR), SB = sum(SB)) 
fg %>% left_join(mas, by = "playerID") %>% 
  filter(HR >= 300 & SB >= 300)%>%
  arrange(HR)
```

#### Or I like to combine above data and get same result in same codu chunks.
```{r}
Batting %>%
  group_by(playerID) %>%
  summarise(HR = sum(HR), SB = sum(SB)) %>%
  filter(HR >= 300 & SB >= 300) %>%
  left_join(Master, by = "playerID") %>%
  select(nameFirst, nameLast, HR, SB)

```
### Second Part
#### Similarly, name every pitcher in baseball history who has accumulated at 
#### least 300 wins (W column) and at least 3,000 strikeouts (SO column).

```{r}

Pitching %>% 
  group_by(playerID) %>% 
  summarize(W = sum(W), SO = sum(SO)) %>%
  filter(W >= 300 & SO >= 3000) %>% 
  left_join(mas, by = "playerID") %>%
  arrange(W)

```
#### Or other way
```{r}
Pitching %>%
  group_by(playerID) %>%
  summarise(W = sum(W), SO = sum(SO)) %>%
  filter(W >= 300 & SO >= 3000) %>%
  left_join(Master, by = "playerID") %>%
  select(nameFirst, nameLast, W, SO)

```

### Third Part
#### Identify the name and year of every player who has hit at least 50 home 
#### runs in a single season. Let table1 refer to the data frame that contains 
#### this information. For each season that appeared in the data frame table1, 
#### find the player that has the lowest batting average that season. 
#### Hint: Use a semi_join

```{r}
mas <- Master %>% 
  group_by(playerID) %>%
  summarize( playerID = paste(playerID),
             name = paste(nameFirst, nameLast, sep = " "))
fg <- Batting %>% 
  group_by( playerID, yearID) %>% 
  select(playerID, yearID, HR,H,AB)
fg1 <- fg %>% 
  semi_join(mas, by = "playerID") %>% 
  filter(HR >= 50)
```

#### first dentify the name and year of every player who has hit at least 50 
#### home runs in a single season
```{r}
fg1 %>% left_join(mas, by = "playerID") %>% 
  arrange(desc(HR))
```

#### find the player that has the lowest batting average that season

```{r}
last <- Batting %>% 
  group_by(playerID, yearID) %>% 
  filter(HR >= 50) %>% 
  mutate(average = sum(H)/sum(AB)) %>% 
  select(playerID, yearID, HR, average) %>%
  arrange(average)

mas %>% 
  right_join(last, by = "playerID") %>% 
  select(HR,average,name,yearID) %>% 
  arrange(average)%>% 
  ungroup() 
```

#### Or Other way we get same result.

```{r}
Batting %>%
  group_by(playerID, yearID) %>%
  summarise(HR = sum(HR), average = sum(H)/sum(AB)) %>%
  filter(HR >= 50) %>%
  left_join(Master, by = "playerID") %>%
  select(nameFirst, nameLast, HR, average) %>%
  ungroup() %>%
  arrange(average)

```


#### THE END