---
title: "Midterm1"
author: "Rabin Paudel"
date: "9/30/2021"
output: html_document
---

### Honor pledge
#### Signature: Rabin Paudel

## Problem I (20pts)

#### Library together with the make_babynames_dist()function from the mdsr libra
#### ry. A snippet of the data is given below.

```{r}
library(babynames)
library(mdsr)
names_df <- make_babynames_dist()
names_df
```
#### The Hmisc library is useful to get the wtd.quantile function to compute the
#### weighted quantile for the age distribution below

```{r}
library(Hmisc)
library(dplyr)
library(tidyr)
names_age_dist <- names_df %>%
filter(n > 100) %>%
group_by(name,sex) %>%
mutate(wgt = est_alive_today / sum(est_alive_today)) %>%
dplyr::summarize(
N = n(),
est_num_alive = sum(est_alive_today),
quantiles = list(
wtd.quantile(
age_today, est_alive_today, probs = c(0.25, 0.5, 0.75), na.rm = TRUE
)
)
) %>%
mutate(measures = list(c("q1_age", "median_age", "q3_age"))) %>%
unnest(cols = c(quantiles, measures)) %>%
pivot_wider(names_from = measures, values_from = quantiles) %>% ungroup()
names_age_dist
```

```{r}
names_max_iqr <- names_age_dist %>% mutate(iqr = q3_age - q1_age) %>%
filter(N >= 100) %>% group_by(sex) %>%
slice_max(iqr,n=10)
names_max_iqr
```


#### Using this data, generate the following similar figure.
```{r}
age_breaks <- head(1:8 * 10 + 0)

p0 <- ggplot(data = names_max_iqr, 
             aes(x = reorder(name, -median_age)), y = median_age)+ xlab(NULL) + 
  ylab(NULL) + ggtitle(~bold("Names With the Widest Age Spreads"),
subtitle = "Estimated interquartile range of ages for Americans alive as of 
Jan.1, 2014") 
p1 <-  p0 + geom_linerange(aes(ymin = q1_age, ymax = q3_age, color = sex),
                           size = 10, alpha = 0.7, show.legend = NULL) 
  
p2 <- p1 + scale_y_continuous(breaks = age_breaks)+
  geom_point(aes(x = 7, y = 85), fill = "#ed3324", 
             color = "white", size = 4, shape = 21) +
  geom_point(aes(y = median_age), fill = "#ed3324", 
             color = "white", size = 4, shape = 21) +
  geom_text(aes(y = 90, x = 7, label = "median")) 
   
p3 <- p2 + facet_wrap(~sex, nrow=2, scales = "free_y") + geom_hline(
  data = tibble(x = age_breaks), 
  aes(yintercept = x), 
  linetype = 3
) +
  coord_flip()
p3

```

## Problem II (20pts)
#### The next problem uses the weather dataframe from the nycflights13 library.

```{r}
library(nycflights13)
library(dplyr)
weather
```
#### Using this data, write code snippets using dplyr functions to answer the 
#### following questions.
### (a) The ten most rainy day in the NYC metro area in 2013.

```{r}
weather %>% group_by(origin, year, month, day) %>% summarise(precip) %>% 
arrange(desc(precip))%>%
head(10)

```

### (b) To find three consecutive days in 2013 where the highest temp. is at 
### least 95 degrees Farenheit.

```{r}
new <- weather %>% filter(origin == "JFK") %>% group_by(year,month,day) %>% 
  summarise(max_temp = max(temp)) %>% ungroup() 
new %>% arrange(year, month, day) %>% filter(lag(max_temp) > 95)
```

### (c) Find the dates and time of the ten observations with the largest
### temperature difference between the three airports

```{r}
library(nycflights13)
library(dplyr)
library(tidyr)
weather_subset <- weather %>% select(origin,year,month,day,hour,temp)
weather_temp_wide <- weather_subset %>% group_by(origin,year,month,day,hour) %>%
dplyr::summarize(temp = mean(temp)) %>%
pivot_wider(names_from = "origin", values_from = "temp") %>% ungroup()
weather_temp_wide
```

#### The largest temperture difference are.
```{r}
weather_temp_wide %>% mutate(temp_diff = pmax(EWR, JFK, LGA) - 
                               pmin(EWR,JFK,LGA)) %>%
  arrange(desc(temp_diff)) %>% head(10)
```

## Problem III(20pts)
#### This problem uses data on song rankings for billboard top 100 in the years
#### 2000. A snippet of the data is given below.

```{r}
library(tidyr)
billboard
```

#### Using this data, write code to convert the data into a long format similar 
#### to the one below.

```{r}
cd <- billboard  %>% pivot_longer(cols = starts_with("wk"), names_to = "week",
                                   values_to = "rank", values_drop_na = TRUE) 
cd %>% mutate(week = as.numeric(gsub("wk", "", week)))
```

## Problem IV(30pts)

#### The data is part of the fec16 library. Please refer to the given link for 
#### more details about the data.

```{r}
## First do an install.packages("fec16") if the fec16 library is missing.
library(fec16)
## If you are using Rmarkdown you might want to run this chunk only once
## or set the chunk option cache=TRUE
## so as to not download the data again and again and again
all_contributions <- read_all_contributions()
all_contributions
```

### a. (10pts)
#### challenger candidate(s) who won against the in-cumbent candidate(s) ? 
#### Please include the name of challenger candidate(s) who won and the 
#### incumbent candidates who lost.

```{r}
newData <- candidates %>% left_join(results_senate, by = "cand_id") %>%
   filter(cand_ici == "C", won == "TRUE") %>% summarise(cand_name)
newData 

```

### b. (10pts)
#### Draw a scatterplot visualizing the relationship between the total amount of
#### contributions a candidate received and the number of votes they got in 
#### the US senate elections

```{r}
result <- left_join(results_senate, candidates, by = "cand_id") %>%
  left_join(all_contributions, by = "cand_id") %>%
  group_by(cand_id, cand_name, cand_pty_affiliation) %>%
  summarise(
    sum_votes = sum(general_votes, na.rm = TRUE),
    contribution = sum(transaction_amt, na.rm = TRUE)
  ) %>% filter(sum_votes>10)
head(result)
result
```
#### Draw a scatterplot and visualizing the data. 
```{r}
ggplot(result) + geom_point(aes(x = contribution, y = sum_votes))+
  labs(
    title = "Contributions vs. Votes Win",
    x = "Contributions in US Dollars", y = " Amount of Votes"
  )
```

### c. (10pts)
#### The type of interest group for each committee is recorded as part of the 
#### org_tp variable in the commitees data frame. For example.

```{r}
library(fec16)
committees %>% select(org_tp) %>% pull %>% table(useNA = "always")
```

#### You might need the candidate id for the two candidates Hillary Clinton 
#### and Donald Trump. You can do so using the following code chunk example.

```{r}
## You might need to first do an install.packages("stringr")
## if the stringr library cannot be loaded.
candidates %>% filter(stringr::str_detect(cand_name, "CLINTON"))
candidates %>% filter(stringr::str_detect(cand_name, "TRUMP")) 
```

#### Only two committees L,C for CLINTON.
```{r}
join <- left_join(contributions, candidates, by = "cand_id") %>%
  filter( cand_id == "P00003392" |cand_id == "P60012333") %>% 
  summarise(cmte_id, transaction_amt, cand_id) %>%
  left_join(committees, by = "cmte_id") %>%
  select(transaction_amt, org_tp) %>% filter(org_tp != "NA") %>%
  arrange(desc(org_tp))
join
```

#### Only one committees M for TRUMP.
```{r}
join1 <- left_join(contributions, candidates, by = "cand_id") %>%
  filter( cand_id == "P60018835" |cand_id == "P60018991" | cand_id == "P80001571") %>% 
  summarise(cmte_id, transaction_amt, cand_id) %>%
  left_join(committees, by = "cmte_id") %>% 
  select(transaction_amt, org_tp) %>% filter(org_tp != "NA") %>%
  arrange(desc(org_tp))
join1
```
### THE END







