---
title: "Assignment4"
author: "Rabin Paudel"
date: "10/18/2021"
output: pdf_document
---
## Problem 1 (30 pts)
#### This problem uses the corpus of common words in the stringr library. 
#### You can load this corpus using the following command.
```{r}
library(stringr) ## First do an install.packages("stringr") if this library is missing.
data (words)
head(words) ## A snippet of the data
```

### 1.a (10 pts)
#### sing the above corpus, create regular expressions that find all words that 
#### (each of the following is a separate regexp)
#### Start with y 
```{r}
words %>% str_subset("^y")
```

#### Or

```{r}
str_view(words, "^y", match = TRUE)
```

#### End with x
```{r}
words %>% str_subset("x$")
```
#### Or
```{r}
str_view(words, "x$", match = TRUE)
```

#### Are exactly six letters long (you are not allowed to use str_length() :-))

```{r}
str_view(words, "^......$", match = TRUE)
```

#### Are exactly five letters long and is a palindrome.

```{r}
str_view(words, "^(.)(.).\\2\\1$", match = TRUE)
```

#### End with ed but not eed.

```{r}
str_view(words, "[^e]ed$", match = TRUE)
```

#### Start with three consonants (the letter y is a consonant) or have three or 
#### more vowels in a row.

```{r}
str_view(words, "^[^aeoiouy]{3}" , match = TRUE) 
str_view(words, "[aeiou]{3}", match = TRUE)
```

#### Have two or more vowel-consonant pairs in a row.

```{r}
str_view(words, "([aeiou][^aeiou]){2,}", match = TRUE)
```

#### Start and end with the same character.

```{r}
str_view(words, "^(.).*\\1$", match = TRUE)
```

#### Contain one letter repeated in at least three places (for example the word 
#### eleven contains three es)

```{r}
str_view(words, "(.).+\\1.+\\1", match = TRUE)
```

### b.(10 pts)
### Using the above corpus, find the top words that have
#### the highest number of vowels

```{r}
library(tidyverse)
vowels <- c("a","e","i","o","u")
vowel_number <- tibble(words = words,
                       n_vowel = str_count(words, vowels))
vowel_number %>% arrange(desc(n_vowel)) %>% head(10)
```

#### the highest proportion of vowels.
```{r}
vowel_pro <- tibble(words = words,
                    n_string = str_length(words),
                    n_vowel = str_count(words, vowels),
                    prop_vol = n_vowel/n_string)
vowel_pro %>% arrange(desc(prop_vol)) %>% head(10)
```
### c. (10 pts)
#### ow switch the first and last letters for each word in the words corpus. 
#### Which of those strings are still words ? Hint: See the last example in 
#### section 14.4.4 of R for data science. See also set operations.

#### I just print top 20.
```{r}
str_replace(words, "(^.)(.*)(.$)", "\\3\\2\\1") %>% head(20)
```

## Problem 2 (30 pts)
#### You can download the data as follows.

```{r}
#install.packages("gutenbergr")
library(gutenbergr)
#install.packages("dplyr")
library(dplyr)
library(ggplot2)
shakespeare_complete <- gutenberg_download(100)
shakespeare_complete
```

#### Using this data, answer the following question.

### a. (10pts)

#### se regular expression to determine the number of speaking lines in all of 
#### the works of William Shakespeare. Here we care only about how many times a 
#### character speaks, not what they say or for how long they speak.

#### #We find number of line Particular character speaks columns n

```{r}
new_one <- shakespeare_complete$text %>%str_subset("[:upper:]\\.$") 
new_two <- data.frame("x" = c(new_one))
#new_two
df <- new_two %>% group_by(x) %>% count(x, sort = TRUE) 
df
#or
sha <- shakespeare_complete %>% 
  mutate(field = str_extract(text,"[:upper:]\\.$")) %>% 
  filter(field != "NA") %>% select(text) %>% count(text, sort = TRUE) 
sha
```

### b. (10pts)
#### Make a bar chart or table displaying the top 30 characters with the 
#### greatest number of lines.

```{r}
new_one <- shakespeare_complete$text %>%str_subset("[:upper:]\\.$") 
new_two <- data.frame("x" = c(new_one))
#new_two
df <- new_two %>% group_by(x) %>% count(x, sort = TRUE) %>% head(30)
df
#or
sha <- shakespeare_complete %>% 
  mutate(field = str_extract(text,"[:upper:]\\.$")) %>% 
  filter(field != "NA") %>% select(text) %>% count(text, sort = TRUE) %>%
  head(30) 
sha

ggplot(sha, aes(x=text, y=n)) + 
  geom_bar(stat = "identity", color = "red", width = 0.5) +
  xlab("Text") + ylab("Total Count") + 
  ggtitle("Bar Chart Displaying The Top 30 Characters With Greatest Number")
```

### c. (10pts)
#### ind all hyphenated words used in the complete works of Shakespeare. 
#### How many hyphenated words are there ? Hint: you might want to use the 
#### function str_extract that is part of the stringr library.

```{r}
sha <- shakespeare_complete %>% 
  mutate(fiel = str_extract(text,"(?:[a-z]+-[a-z]+)")) %>% 
  filter(fiel != "NA") %>% select(text) %>% count(text, sort = TRUE)
sha
sha%>% count()
```

