
#### We can instill this library and load the dataset using the 
#### folowing code chunk.
```{r}
#install.packages("gapminder")
#install.packages("dplyr")
library(gapminder)
library(dplyr)
gapminder

```
## Problem 1
## Q.N1 Code 

```{r}
gapminder %>% count(continent)
```
#### Here is the result of number of observation per continent.
#### Africa:   624
#### Americas: 300
#### Asia:     396
#### Europe:   360
#### Oceania:   24

## Q.N2 Code 

```{r}
gapminder  %>% group_by(continent) %>% summarise(CountryCount = 
                                                    n_distinct(country))
```

#### Here is (distinct) countries do we have each continent
#### Africa:   52
#### Americas: 25
#### Asia:     33
#### Europe:   30
#### Oceania:   2

## Q.N3 Code
#### Try to create new column named gdp_ration to find the with the help of 
#### given code.

```{r}
gapminder %>% arrange(country,year) %>% mutate(gdpRatio 
                          = gdpPercap/gdpPercap[country == "United States"])
GDP1 <- mutate (gapminder, GDP = pop*gdpPercap)
GDP1 %>% arrange(country,year) %>% mutate(gdp_ratio = 
                                            GDP/GDP[country == "United States"])

```

## Q.N4 Code
#### Try to find lowest and highest lifeExpectancy.

```{r}
gapminder %>% group_by(country) %>% filter(lifeExp < 30)
gapminder %>% group_by(country) %>% filter(lifeExp > 82)
```
#### The Lowest life expectancy country is: Rwanda in Africa
#### and Afghanistan in Asia.
#### And the highest Life Expectancy country is Japan and Hong Kong in Asia.

## Q.N5 Code
#### To find the country that experienced the sharpest 5 year deop in life
#### expectancy during the period from 1952 to 1997.

```{r}
gapminder %>% group_by(country) %>% mutate(lifeExp_change = 
                                               lifeExp - lag(lifeExp)) %>%
 select(lifeExp_change, everything()) %>% filter(year >= 1952 & 
                                                     year <= 1997 ) %>% 
   filter(continent =="Africa") %>% filter(lifeExp_change < -10) 
```
#### Change continent and Life Expectancy change is every time and enter manually 
#### The Sharpest 5 year drop in life expectancy continent are: 
#### Africa:   Rwanda
#### Americas: El Salvador
#### Asia:     Cambodia
#### Europe:   Bulgaria
#### Oceania:   Non Of the country

## Problem 2
## Q.N1 Code 
```{r}
#install.packages("nycflights13")
library(nycflights13)
flights %>% select(dep_time, sched_dep_time, dep_delay, arr_time, arr_delay)
flights %>% filter( arr_delay > 900 & dep_delay >900) %>% 
    select(dep_delay, tailnum)

flights %>% group_by(tailnum) %>% filter(tailnum == "N384HA") %>%
      filter(arr_delay > 0 & dep_delay) %>% 
      summarise(NumOfFlight = n_distinct(tailnum))
```

#### The Flight N384HA worst flight one time record total number of delay is 1301.

## Q.N2 Code 
#### Yes, the proportion of canceled flights per day related to the average 
#### delay per day.

```{r}
flights %>% group_by(dest) %>%
   filter(n() > 365, arr_delay >0) %>%
   mutate(prop_canflight = arr_delay/sum(arr_delay)) %>%
   select(year:day, dest, arr_delay, prop_canflight)
```

## Q.N3 Code

```{r}
library(dplyr)
 library(nycflights13)
 data(flights)
 flights_tod <- flights %>% mutate(time_of_day = case_when(
   sched_dep_time <= 1100 ~ "morning",
   between(sched_dep_time,1101,1400) ~ "noon",
   between(sched_dep_time,1401,1800) ~ "afternoon",
   sched_dep_time > 1801 ~ "evening"
 )) %>% select(time_of_day, everything())
 flights_tod
flights_tod %>% filter(time_of_day == "morning" & arr_delay > 0) %>%  
  summarise(MorningDelay = n_distinct(arr_delay))
flights_tod %>% filter(time_of_day == "noon" & arr_delay > 0) %>%  
  summarise(noonDelay = n_distinct(arr_delay))  
flights_tod %>% filter(time_of_day == "afternoon" & arr_delay > 0) %>%  
  summarise(afternoonDelay = n_distinct(arr_delay))   
flights_tod %>% filter(time_of_day == "evening" & arr_delay > 0) %>%  
  summarise(eveningDelay = n_distinct(arr_delay))  
```

#### noon flight better because evening time less delay then other time.

## Q.N4 Code
#arrange the flights in order of month followed by day followed by 
#deptime before using lag.

```{r}
flights_new <- flights %>% arrange(month,day,dep_time) %>% group_by(origin) %>% 
mutate(delay = lag(dep_delay))
flights_new
flights_new <- flights %>% arrange(month,day,dep_time) %>% group_by(origin) %>% 
   mutate(prev_dep_delay = lag(dep_delay)) %>% ungroup()
flights_new
```

## The End